---
title: "Log Odds"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r import, child=c('../data/import.Rmd','../wk03/plots02.Rmd')}

```

## Log-Odds
```{r log_odds transform}
# log-odds by interest trait: religion, sex, ideology, role, education
# log-odds grouped by religion
religion_lo = fulltidy %>%
  filter(!is.na(religion)) %>%
  count(Word, religion, sort = TRUE) %>%
  bind_log_odds(religion, Word, n) %>%
  arrange(-log_odds_weighted) %>%
  group_by(Word) %>%     ## only keep words used over 50 times total
  filter(sum(n) > 10) %>% # how robust is log-odds
  ungroup %>%
  group_by(religion) %>%
  slice_max(log_odds_weighted, n = 10)    ## most likely by religion

# log-odds grouped by sex
sex_lo = fulltidy %>%
  filter(!is.na(sex)) %>%
  count(Word, sex, sort = TRUE) %>%
  bind_log_odds(sex, Word, n) %>%
  arrange(-log_odds_weighted) %>%
  group_by(Word) %>%     ## only keep words used over 50 times total
  filter(sum(n) > 10) %>% # how robust is log-odds
  ungroup %>%
  group_by(sex) %>%
  slice_max(log_odds_weighted, n = 10)    ## most likely by sex

# log-odds grouped by ideology
ideology_lo = fulltidy %>%
  filter(!is.na(ideology)) %>%
  count(Word, ideology, sort = TRUE) %>%
  bind_log_odds(ideology, Word, n) %>%
  arrange(-log_odds_weighted) %>%
  group_by(Word) %>%     ## only keep words used over 50 times total
  filter(sum(n) > 10) %>% # how robust is log-odds
  ungroup %>%
  group_by(ideology) %>%
  slice_max(log_odds_weighted, n = 10)    ## most likely by ideology

# log-odds grouped by role
role_lo = fulltidy %>%
  filter(!is.na(Role)) %>%
  count(Word, Role, sort = TRUE) %>%
  bind_log_odds(Role, Word, n) %>%
  arrange(-log_odds_weighted) %>%
  group_by(Word) %>%     ## only keep words used over 50 times total
  filter(sum(n) > 10) %>% # how robust is log-odds
  ungroup %>%
  group_by(Role) %>%
  slice_max(log_odds_weighted, n = 10)    ## most likely by role

# log-odds grouped by role
edu_lo = fulltidy %>%
  filter(!is.na(edu)) %>%
  count(Word, edu, sort = TRUE) %>%
  bind_log_odds(edu, Word, n) %>%
  arrange(-log_odds_weighted) %>%
  group_by(Word) %>%     ## only keep words used over 50 times total
  filter(sum(n) > 10) %>% # how robust is log-odds
  ungroup %>%
  group_by(edu) %>%
  slice_max(log_odds_weighted, n = 10) ## most likely by education
```
Log Odds is a Bayesian technique comparing the frequency of the words in the text to the frequency of the words in each grouping. This technique differs from term frequency-inverse document frequency, because  tf-idf uses the frequency of the words and compares which words are different across documents, using a relative search to estimate importance. Log-odds approach is different, using [what is log-odds] to weight variable 

```{r log_odds relgion plot}
##### plots #####
religion_lo %>%
  ggplot(aes(log_odds_weighted, fct_reorder(Word, log_odds_weighted),fill = religion)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~religion, scales = "free_y") +
  labs(y = NULL)
```
Using log-odds to examine religion, we can learn a lot more about how people 

```{r log_odds sex}
sex_lo %>%
  ggplot(aes(log_odds_weighted, fct_reorder(Word, log_odds_weighted),fill = sex)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sex, scales = "free") +
  labs(y = NULL)
```
For sex, we can see that [observations]

```{r log_odds ideology}
ideology_lo %>%
  ggplot(aes(log_odds_weighted, fct_reorder(Word, log_odds_weighted),fill = ideology)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ideology, scales = "free_y") +
  labs(y = NULL)
```
With respect to political ideology, [observations]

```{r log_odds role}
role_lo %>%
  ggplot(aes(log_odds_weighted, fct_reorder(Word, log_odds_weighted), fill = Role)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~Role, scales = "free_y") +
  labs(y = NULL)
```
[observations] These plots look completely different, the persuaders tending towards specific, large words and nearly all of the persuadees using positively associated filler words, like 'oh', 'ok', 'sure'

```{r log_odds edu}
edu_lo %>%
  ggplot(aes(log_odds_weighted, fct_reorder(Word, log_odds_weighted), fill = edu)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~edu, scales = "free_y") +
  labs(y = NULL)
```
Most words have a fairly uniform frequency within the groups, but it appears the postgraduates kept closely to the top words from the overall sample. Less than college found 'iam' and 'ur', shorthand for I am and you are. 'mexico' also made a comeback (liberal/atheist) and 'iam'/'ur' (moderate)
