---
title: "Taking A Look"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# keeping track of the code that each script requires! Reproducible
```

```{r import, child='../data/import.Rmd'}

```

## Exploratory Analysis

What's in our data? Let's take a look through some plots.

```{r fulltidy plots}
# looking at individual sentences/instances to look at how standard the data are
fullset %>% slice_sample(n = 5) %>% pull(Dialogue)
# most popular words
fulltidy %>% count(Word, sort = TRUE) %>% 
  filter(n > 1000) %>% 
  ggplot(aes(fct_reorder(Word, n),n)) +
  geom_col() +
  coord_flip() +
  labs(x = NULL)
```
Our data are held as full exchanges in a dialogue, where each exchange can comprise of multiple sentences. This plot shows us which words were used the most during the dialogues across the whole set, including a number of words such as 'can' and 'like', words that could be filtered our due to their lack of inherent meeting. The words 'children','save',and 'donate' were used the most out of any words, most likely because the charity is called 'Save the Children' and the participants are tasked with convincing people to donate.

```{r fulltidy plots2}
# zipf's law = nl follows log scale, log-log plot w/ rank?
fulltidy %>% count(Word, sort = TRUE) %>% 
  # filter(n > 500) %>% 
  slice_max(n,n = 25) %>% 
  ggplot(aes(Word,n)) +
  geom_col() +
  coord_flip() +
  labs(title = 'Top Used Words in Dialogue', x = NULL)

fulltidy %>% count(Word,Role, sort = TRUE) %>% 
  # filter(n > 500) %>% 
  slice_max(n,n = 25) %>% 
  ggplot(aes(Word,n, fill = as.factor(Role))) +
  geom_col() +
  coord_flip() +
  labs(title = "Top Words by Role", x = NULL)

####Bar Charts (n) By Role####

fulltidy %>% 
  filter(Role == 1) %>% 
  count(Word, sort = TRUE) %>% 
  slice_max(n,n = 20) %>% 
  ggplot(aes(Word,n)) +
  geom_col(aes(fill = n)) +
#  geom_text(aes(label = Word), check_overlap = TRUE, vjust = 1.5) +
  labs(x = NULL, title = "Words: Persuadee") +
  coord_flip()

fulltidy %>% 
  filter(Role == 0) %>% 
  count(Word, sort = TRUE) %>% 
  slice_max(n,n = 20) %>% 
  ggplot(aes(Word,n)) +
  geom_col(aes(fill = n)) +
#  geom_text(aes(label = Word), check_overlap = TRUE, vjust = 1.5) +
  labs(x = NULL, title = "Words: Persuader") +
  coord_flip()

#####Bar Charts (n/sum(n) by Role####
# proportional, persuadee bar
# fulltidy %>% 
#   filter(Role == 1) %>% 
#   count(Word, sort = TRUE) %>% 
#   mutate(proportion = n / sum(n)) %>% 
#   ggplot(aes(Word,proportion)) +
#   geom_histogram(aes(color = n)) +
#   geom_text(aes(label = Word), check_overlap = TRUE, vjust = 1.5) +
#   labs(x = NULL, title = "Words: Persuadee") +
#   coord_flip()
# 
# # proportional, persuader bar
# fulltidy %>% 
#   filter(Role == 0) %>% 
#   count(Word, sort = TRUE) %>% 
#   mutate(proportion = n / sum(n)) %>% 
#   ggplot(aes(Word,proportion)) +
#   geom_jitter(aes(color = n)) +
#   geom_text(aes(label = Word), check_overlap = TRUE, vjust = 1.5) +
#   labs(x = NULL, title = "Words: Persuader") +
#   coord_flip()

```

## Including Plots
```{r annset plots, echo=FALSE}
# plot multiple things against actual donation
# plot of actual donation v stated donation
# best way to represent the donation amount and the intended donation amount
annset %>% 
#  filter(IntDonation < 5000) %>% 
  ggplot() +
  geom_boxplot(aes(IntDonation))
annset %>%
  filter(IntDonation < 75) %>% 
  ggplot() +
  geom_boxplot(aes(Donation-IntDonation))

annset %>% 
  ggplot(aes(IntDonation,Donation)) +
  geom_jitter()

annset %>% 
  #difference between IntDonation and Donation
  filter(IntDonation < 75) %>% 
  ggplot(aes(Donation-IntDonation)) +
  geom_histogram(binwidth = 10) 


annset %>% 
  #difference between IntDonation and Donation
  filter(IntDonation < 75) %>% 
  ggplot(aes(Donation-IntDonation)) +
  geom_histogram(bins = 15) 

```
Most people actually donated how much they pledged to donate. As expected, some donated less. There are two outliers with respect to x, one where someone promise to donate \$10,000 but didn't, and another where someone promised to donate \$500 and didn't. A kind soul donated more than they initially pledged. 

```{r fullset plots, echo=FALSE}
# distributions for each personality trait (one of those bar plots that is stacked/filled)

# utterances per conversation
fullset %>% 
  count(ConvoId) %>% 
  ggplot(aes(n)) +
  geom_histogram()
##### values plots #####
# the big five [extrovert, agreeable, conscientious, neurotic, open]
full_i_raw %>% 
  select(B3,extrovert:open) %>% 
  pivot_longer(extrovert:open,names_to = "trait", values_to = "score") %>% 
  ggplot(aes(score, fill = trait)) +
  facet_wrap(~trait) +
  geom_histogram(bins = 5, show.legend = FALSE)

# the moral foundations [care, fairness, loyalty, authority, purity, freedom]
full_i_raw %>% 
  select(B3,care:freedom) %>% 
  pivot_longer(care:freedom,names_to = "trait", values_to = "score") %>% 
  ggplot(aes(score, fill = trait)) +
  facet_wrap(~trait) +
  geom_histogram(bins = 6, show.legend = FALSE)

# the Schwartz Portrait [conform, tradition, benevolence, universalism, self_direction, stimulation, hedonism, achievement, power, security]
full_i_raw %>% 
  select(B3,conform:security) %>% 
  pivot_longer(conform:security,names_to = "trait", values_to = "score") %>% 
  ggplot(aes(score, fill = trait)) +
  facet_wrap(~trait) +
  geom_histogram(bins = 6, show.legend = FALSE)

# the decision making style [rational, intuitive]
full_i_raw %>% 
  select(B3,rational:intuitive) %>% 
  pivot_longer(rational:intuitive,names_to = "trait", values_to = "score") %>% 
  ggplot(aes(score, fill = trait)) +
  facet_wrap(~trait) +
  geom_histogram(bins = 5, show.legend = FALSE)

#####comments####
# # 10 baby plots of value plots
# full_i_raw %>% 
#   select(B3,fairness:self_direction) %>% 
#   pivot_longer(fairness:self_direction,names_to = "trait", values_to = "score") %>% 
#   ggplot(aes(score, fill = trait)) +
#   facet_wrap(~trait) +
#   geom_histogram(bins = 6, show.legend = FALSE)

# # 3 barcharts for education/ideology
# full_i_raw %>% 
#   count(edu,ideology) %>% 
#   mutate(ideology = factor(ideology,c("Conservative","Moderate","Liberal"))) %>% 
#   # reoder education levels
#   filter(!is.na(edu), !is.na(ideology)) %>% 
#   ggplot(aes(ideology,n,fill = ideology)) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~edu,scales = "free")

##### traits plots ####
# one (or more) trait plots with other info
# more exploratory plots age <dbl>, sex <chr>, race <chr>, edu <chr>, marital <chr>, employment <chr>, income <dbl>, religion <chr>, ideology <chr>

# age/sex
full_i_raw %>%
  filter(!is.na(sex)) %>% 
  ggplot(aes(age)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)
# income/employment
full_i_raw %>%
  filter(!is.na(employment),!is.na(income)) %>% 
  ggplot(aes(income,fill = marital)) +
  geom_histogram(binwidth = 1.5) +
  facet_wrap(~employment) +
  coord_flip()

# religion/ideology
full_i_raw %>%
  count(religion,ideology) %>% 
  filter(!is.na(ideology),!is.na(religion)) %>% 
  ggplot(aes(ideology,n)) +
  geom_col() +
  facet_wrap(~religion)

# # stacked barchart education/ideology
# full_i_raw %>% 
#   count(edu,ideology) %>% 
#   mutate(ideology = factor(ideology,c("Conservative","Moderate","Liberal"))) %>% 
#   # reorder education levels
#   filter(!is.na(edu), !is.na(ideology)) %>% 
#   ggplot(aes(ideology,n,fill = edu)) +
#   geom_col()

# dodge barchart ideology/education
full_i_raw %>% 
  count(edu,ideology) %>% 
  mutate(ideology = factor(ideology, c("Conservative","Moderate","Liberal"))) %>% 
  mutate(edu = factor(edu,c("Less than four-year college","Four-year college","Postgraduate"))) %>% 
  # reoder education levels
  filter(!is.na(edu), !is.na(ideology)) %>% 
  ggplot(aes(ideology,n,fill = edu)) +
  geom_col(position = "dodge")

# # dodge barchart religion/ideology
# full_i_raw %>% 
#   count(religion,ideology) %>% 
#   mutate(ideology = factor(ideology,c("Conservative","Moderate","Liberal"))) %>% 
#   # reoder education levels
#   filter(!is.na(religion), !is.na(ideology)) %>% 
#   ggplot(aes(religion,n,fill = ideology)) +
#   geom_col(position = "dodge")

# dodge barchart ideology/religion
full_i_raw %>% 
  count(religion,ideology) %>% 
  mutate(ideology = factor(ideology,c("Conservative","Moderate","Liberal"))) %>% 
  mutate(religion = factor(religion,c("Catholic","Protestant","Atheist","Other religion"))) %>% 
  # reoder education levels
  filter(!is.na(religion), !is.na(ideology)) %>% 
  ggplot(aes(ideology,n,fill = religion)) +
  geom_col(position = "dodge")

```


```{r other plots}
# plots for frequency of (top) words and donation amount
# have to use the id_tdf words to organize by top words that ISN'T a stop word


# top words by personality trait and role
# have to use the id_tdf words to organize by top words that ISN'T a stop word

```




