---
title: "Taking A Look"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# keeping track of the code that each script requires! Reproducible
```

```{r import, child='../data/import.Rmd'}

```

## Text Analysis

Using tidy data, we can explore text data and unsupervised learning to gain a better understanding of our dataset.
For this project, we are using MIT's Persuasion for Good dataset </link here>

```{r fulltidy plots 01}
# looking at individual sentences/instances to look at how standard the data are
fullset %>% slice_sample(n = 5) %>% pull(Dialogue)
# most popular words
fulltidy %>% count(Word, sort = TRUE) %>% 
  filter(n > 1000) %>% 
  # reorder words by count
  ggplot(aes(fct_reorder(Word, n),n)) +
  geom_col() +
  coord_flip() +
  labs(x = NULL, title = "Most Common Words in Dialogue")
```
Our data are held as full exchanges in a dialogue, where each exchange can comprise of typed exchanges between participants.
Using 'tidy' data, we can see which words were used the most during the all the conversations, including a number of words such as 'can' and 'like', words that could be filtered our due to their lack of inherent meeting. The words 'children','save',and 'donate' were used the most out of any words, most likely because the charity is called 'Save the Children' and the participants are tasked with convincing people to donate.

```{r fulltidy plots 02/ dont print, eval=FALSE}
# zipf's law = nl follows log scale, log-log plot w/ rank?
fulltidy %>% count(Word, sort = TRUE) %>% 
  # filter(n > 500) %>% 
  slice_max(n,n = 25) %>% 
  ggplot(aes(fct_reorder(Word,n),n,)) +
  geom_col() +
  coord_flip() +
  labs(title = 'Top Used Words in Dialogue', x = NULL)
```


```{r fulltidy plots 03}
fulltidy %>% count(Word,Role, sort = TRUE) %>% 
  # filter(n > 500) %>% 
  slice_max(n,n = 25) %>% 
  ggplot(aes(Word,n, fill = as.factor(Role))) +
  geom_col() +
  coord_flip() +
  labs(title = "Top Words by Role", x = NULL, fill = NULL)
```
One of the major facets we examine is the role in the conversation. Each person is assigned a role, the person asking for a donation and the person being asked to donate.

```{r fulltidy plots 04}
####Bar Charts (n) By Role####

fulltidy %>% 
  count(Word,Role,sort = TRUE) %>% 
  slice_max(n,n = 20) %>% 
  ggplot(aes(Word,n)) +
  geom_col(aes(fill = n),show.legend = FALSE) +
#  geom_text(aes(label = Word), check_overlap = TRUE, vjust = 1.5) +
  facet_wrap(~Role, scales = "free") +
  labs(x = NULL, y = 'count',title = "Words by Role") +
  coord_flip()
```
This representation shows the most common words by Role. We'll come back 

```{r fulltidy plots comment, comment=FALSE}
# fulltidy %>% 
#   filter(Role == 0) %>% 
#   count(Word, sort = TRUE) %>% 
#   slice_max(n,n = 20) %>% 
#   ggplot(aes(Word,n)) +
#   geom_col(aes(fill = n)) +
# #  geom_text(aes(label = Word), check_overlap = TRUE, vjust = 1.5) +
#   labs(x = NULL, title = "Words: Persuader") +
#   coord_flip()

#####Bar Charts (n/sum(n) by Role####
# proportional, persuadee bar
# fulltidy %>% 
#   filter(Role == 1) %>% 
#   count(Word, sort = TRUE) %>% 
#   mutate(proportion = n / sum(n)) %>% 
#   ggplot(aes(Word,proportion)) +
#   geom_histogram(aes(color = n)) +
#   geom_text(aes(label = Word), check_overlap = TRUE, vjust = 1.5) +
#   labs(x = NULL, title = "Words: Persuadee") +
#   coord_flip()
# 
# # proportional, persuader bar
# fulltidy %>% 
#   filter(Role == 0) %>% 
#   count(Word, sort = TRUE) %>% 
#   mutate(proportion = n / sum(n)) %>% 
#   ggplot(aes(Word,proportion)) +
#   geom_jitter(aes(color = n)) +
#   geom_text(aes(label = Word), check_overlap = TRUE, vjust = 1.5) +
#   labs(x = NULL, title = "Words: Persuader") +
#   coord_flip()

```
Text analysis is based around frequency, how often the words occur. And there is something meaningful to learn from the different methods of exploring frequency we go over in this analysis. We'll go through

## Donations Aside
```{r annset plots, echo=FALSE}
# plot multiple things against actual donation
# plot of actual donation v stated donation
# best way to represent the donation amount and the intended donation amount
annset %>% 
#  filter(IntDonation < 5000) %>% 
  ggplot() +
  geom_boxplot(aes(IntDonation))
annset %>%
  filter(IntDonation < 75) %>% 
  ggplot() +
  geom_boxplot(aes(Donation-IntDonation))

annset %>% 
  ggplot(aes(IntDonation,Donation)) +
  geom_jitter()

annset %>% 
  #difference between IntDonation and Donation
  filter(IntDonation < 75) %>% 
  ggplot(aes(Donation-IntDonation)) +
  geom_histogram(binwidth = 10) 


annset %>% 
  #difference between IntDonation and Donation
  filter(IntDonation < 75) %>% 
  ggplot(aes(Donation-IntDonation)) +
  geom_histogram(bins = 15) 

```
Most people actually donated how much they pledged to donate. As expected, some donated less. There are two outliers with respect to x, one where someone promise to donate \$10,000 but didn't, and another where someone promised to donate \$500 and didn't. A kind soul donated more than they initially pledged. 

```{r fullset_plots, echo=FALSE}
# distributions for each personality trait (one of those bar plots that is stacked/filled)

# utterances per conversation
fullset %>% 
  count(ConvoId) %>% 
  ggplot(aes(n)) +
  geom_histogram()
##### values plots #####
# the big five [extrovert, agreeable, conscientious, neurotic, open]
full_i_raw %>% 
  select(B3,extrovert:open) %>% 
  pivot_longer(extrovert:open,names_to = "trait", values_to = "score") %>% 
  ggplot(aes(score, fill = trait)) +
  facet_wrap(~trait) +
  geom_histogram(bins = 5, show.legend = FALSE) +
  labs(title = 'the big five')
```
Here we have our distributions of personality traits. This study collected data for three major personality scores: the big five, the Schwartz and the moral foundations. The big five show us that our sample has very agreeable, extroverted, and conscientious participants. They also skew towards open. 

```{r fullset plots, echo=FALSE}
# the moral foundations [care, fairness, loyalty, authority, purity, freedom]
full_i_raw %>% 
  select(B3,care:freedom) %>% 
  pivot_longer(care:freedom,names_to = "trait", values_to = "score") %>% 
  ggplot(aes(score, fill = trait)) +
  facet_wrap(~trait) +
  geom_histogram(bins = 6, show.legend = FALSE)
```
Here we see high levels of fairness and care, indicating a compassionate bunch.

```{r fullset plots, echo=FALSE}
# the Schwartz Portrait [conform, tradition, benevolence, universalism, self_direction, stimulation, hedonism, achievement, power, security]
full_i_raw %>% 
  select(B3,conform:security) %>% 
  pivot_longer(conform:security,names_to = "trait", values_to = "score") %>% 
  ggplot(aes(score, fill = trait)) +
  facet_wrap(~trait) +
  geom_histogram(bins = 6, show.legend = FALSE)
```
The Schwartz portrait shows benevolence, self-direction, and universalism

```{r fullset plots, echo=FALSE}
# the decision making style [rational, intuitive]
full_i_raw %>% 
  select(B3,rational:intuitive) %>% 
  pivot_longer(rational:intuitive,names_to = "trait", values_to = "score") %>% 
  ggplot(aes(score, fill = trait)) +
  facet_wrap(~trait) +
  geom_histogram(bins = 5, show.legend = FALSE)

 #####comments####
# # 10 baby plots of value plots
# full_i_raw %>% 
#   select(B3,fairness:self_direction) %>% 
#   pivot_longer(fairness:self_direction,names_to = "trait", values_to = "score") %>% 
#   ggplot(aes(score, fill = trait)) +
#   facet_wrap(~trait) +
#   geom_histogram(bins = 6, show.legend = FALSE)

# # 3 barcharts for education/ideology
# full_i_raw %>% 
#   count(edu,ideology) %>% 
#   mutate(ideology = factor(ideology,c("Conservative","Moderate","Liberal"))) %>% 
#   # reoder education levels
#   filter(!is.na(edu), !is.na(ideology)) %>% 
#   ggplot(aes(ideology,n,fill = ideology)) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~edu,scales = "free")
```

Other relevant metrics for traits: 
```{r fullset plots, echo=FALSE}
##### traits plots ####
# one (or more) trait plots with other info
# more exploratory plots age <dbl>, sex <chr>, race <chr>, edu <chr>, marital <chr>, employment <chr>, income <dbl>, religion <chr>, ideology <chr>

# age/sex
full_i_raw %>%
  filter(!is.na(sex)) %>% 
  ggplot(aes(age)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)
```


```{r fullset plots, echo=FALSE}
# income/employment
full_i_raw %>%
  filter(!is.na(employment),!is.na(income)) %>% 
  ggplot(aes(income,fill = marital)) +
  geom_histogram(binwidth = 1.5) +
  facet_wrap(~employment) +
  coord_flip()
```


```{r fullset plots, echo=FALSE}
# religion/ideology
full_i_raw %>%
  count(religion,ideology) %>% 
  filter(!is.na(ideology),!is.na(religion)) %>% 
  ggplot(aes(ideology,n)) +
  geom_col() +
  facet_wrap(~religion)
```


```{r fullset plots, echo=FALSE}
# # stacked barchart education/ideology
# full_i_raw %>% 
#   count(edu,ideology) %>% 
#   mutate(ideology = factor(ideology,c("Conservative","Moderate","Liberal"))) %>% 
#   # reorder education levels
#   filter(!is.na(edu), !is.na(ideology)) %>% 
#   ggplot(aes(ideology,n,fill = edu)) +
#   geom_col()

# dodge barchart ideology/education
full_i_raw %>% 
  count(edu,ideology) %>% 
  mutate(ideology = factor(ideology, c("Conservative","Moderate","Liberal"))) %>% 
  mutate(edu = factor(edu,c("Less than four-year college","Four-year college","Postgraduate"))) %>%
  # reoder education levels
  filter(!is.na(edu), !is.na(ideology)) %>% 
  ggplot(aes(ideology,n,fill = edu)) +
  geom_col(position = "dodge")

full_i_raw %>% 
  count(edu,ideology) %>% 
  mutate(ideology = factor(ideology, c("Conservative","Moderate","Liberal"))) %>% 
  mutate(edu = factor(edu,c("Less than four-year college","Four-year college","Postgraduate"))) %>%
  # reoder education levels
  filter(!is.na(edu), !is.na(ideology)) %>% 
  ggplot(aes(ideology,n)) +
  geom_col() +
  facet_wrap(~edu)


# # dodge barchart religion/ideology
# full_i_raw %>% 
#   count(religion,ideology) %>% 
#   mutate(ideology = factor(ideology,c("Conservative","Moderate","Liberal"))) %>% 
#   # reoder education levels
#   filter(!is.na(religion), !is.na(ideology)) %>% 
#   ggplot(aes(religion,n,fill = ideology)) +
#   geom_col(position = "dodge")

# dodge barchart ideology/religion
full_i_raw %>% 
  count(religion,ideology) %>% 
  # mutate(ideology = factor(ideology,c("Conservative","Moderate","Liberal"))) %>%
  # mutate(religion = factor(religion,c("Catholic","Protestant","Atheist","Other religion"))) %>%
  # reoder education levels
  filter(!is.na(religion), !is.na(ideology)) %>% 
  ggplot(aes(ideology,n,fill = religion)) +
  geom_col(position = "dodge")

full_i_raw %>% 
  count(religion,ideology) %>% 
  # mutate(ideology = factor(ideology,c("Conservative","Moderate","Liberal"))) %>%
  # mutate(religion = factor(religion,c("Catholic","Protestant","Atheist","Other religion"))) %>%
  # reoder education levels
  filter(!is.na(religion), !is.na(ideology)) %>% 
  ggplot(aes(ideology,n,fill = religion)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~religion)


```


```{r other plots}
# plots for frequency of (top) words and donation amount
# have to use the id_tdf words to organize by top words that ISN'T a stop word


# top words by personality trait and role
# have to use the id_tdf words to organize by top words that ISN'T a stop word

```




